Erosion of somatic tissue identity with loss of the X-linked intellectual disability factor KDM5C

### 2024-07-01 submission/manuscript.Rmd
update chipseq in text counts

```{r, KDM5C_ChIPseq, results='hide', echo = FALSE, warning=FALSE,message=FALSE,error=FALSE}
KDM5C_ChIP_EpiLC <- read.csv("../data/processed/KDM5C_ChIPseq_boundpromoters_EpiLC.csv", header = T)

KDM5C_ChIP_PNC <- read.csv("../data/processed/KDM5C_ChIPseq_boundpromoters_PNC.csv", header = T)

@@

To better elucidate the mechanism behind KDM5C-mediated germline gene silencing, we analyzed KDM5C chromatin immunoprecipitation followed by DNA sequencing (ChIP-seq) datasets we previously generated in EpiLCs[@samantaActivationXistEvolutionarily2022] and primary hippocampal and cortical neuron cultures (PNCs)[@iwaseMouseModelXlinked2016]. We identified a higher number of KDM5C peaks in EpiLCs overall (EpiLCs: `r nrow(KDM5C_ChIP_EpiLC)`, PNCs: `r nrow(KDM5C_ChIP_PNC)`, MACS2 q < 0.1 and fold enrichment > 1) and while KDM5C was primarily localized to gene promoters in both cell types (EpiLCs: , PNCs: , +/- kb TSS), PNCs showed increased localization to distal intergenic regions (EpiLCs: , PNCs: ) (Figure 4A). Gene ontology analysis of KDM5C-bound promoters in EpiLCs revealed high enrichment of germline-related ontologies, including meiotic nuclear division (GO:0140013) and synaptonemal complex assembly (GO:0007130) (Figure 4B). Contrastingly, germline ontologies are not enriched in PNCs and KDM5C-bound genes instead perform non-germline-specific functions, such as heterocycle synthesis (GO: 0018130) and pyrimidine metabolism (GO: 0072527).

>>>>

```{r, KDM5C_ChIPseq, results='hide', echo = FALSE, warning=FALSE,message=FALSE,error=FALSE}
KDM5C_ChIP_EpiLC_promo <- read.csv("../data/processed/KDM5C_ChIPseq_boundpromoters_EpiLC.csv")

KDM5C_ChIP_PNC_promo <- read.csv("../data/processed/KDM5C_ChIPseq_boundpromoters_PNC.csv")

KDM5C_ChIP_EpiLC <- read.csv("../data/raw/ChIPseq_mm10_EpiLC_WTnoKO_consensus_peaks.bed", sep = "\t")

KDM5C_ChIP_PNC <- read.csv("../data/raw/ChIPseq_mm10_PNC_WTnoKO_consensus_peaks.bed", sep = "\t")

@@

To better elucidate the mechanism behind KDM5C-mediated germline gene silencing, we analyzed KDM5C chromatin immunoprecipitation followed by DNA sequencing (ChIP-seq) datasets we previously generated in EpiLCs[@samantaActivationXistEvolutionarily2022] and primary hippocampal and cortical neuron cultures (PNCs)[@iwaseMouseModelXlinked2016]. We identified a higher number of KDM5C peaks in EpiLCs overall (EpiLCs: `r nrow(KDM5C_ChIP_EpiLC)`, PNCs: `r nrow(KDM5C_ChIP_PNC)`, MACS2 q < 0.1 and fold enrichment > 1) and while KDM5C was primarily localized to gene promoters in both cell types (EpiLCs: `r nrow(KDM5C_ChIP_EpiLC_promo)`, PNCs: `r nrow(KDM5C_ChIP_PNC_promo)` +/- `r promorange`kb from TSS), PNCs showed increased localization to distal intergenic regions (EpiLCs: , PNCs: ) (Figure 4A). Gene ontology analysis of KDM5C-bound promoters in EpiLCs revealed high enrichment of germline-related ontologies, including meiotic nuclear division (GO:0140013) and synaptonemal complex assembly (GO:0007130) (Figure 4B). Contrastingly, germline ontologies are not enriched in PNCs and KDM5C-bound genes instead perform non-germline-specific functions, such as heterocycle synthesis (GO: 0018130) and pyrimidine metabolism (GO: 0072527).
